### 2023-03-23 reachable-resilience.tex
up

\problemx{$k$-resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
{ $\Bad \cap \post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
%

%Analogously, we formulate the resilience problem for WSTSs.

\problemx{bounded Resilience problem for WSTS}
{A state $s$ of a WSTS $W=(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
{$\exists k \geq 0$ such that $W$ is $k$-resilient. ?\newline}

\problemx{Resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
{$ ~ \forall s' \in D ~ (s \rightarrow^* s') \implies \exists s'' \in U ~ s' \rightarrow^{*} s''$ ?\newline}

@@

\begin{remark} 
We require $U$ to be upward-closed, as, in the general case if we lift restrictions on $U$ and $D$, then reachability reduces itself to resilience. Indeed, the particular resilience problem where we take $D = \{s\}$ and $U = \{x\}$ is reachability of $x$ from $s$, and thus generally undecidable for WSTS. 
\end{remark}

@@

\begin{proposition}\label{reductions}
coverability and $t$-liveness are reduceable to resilience.
%	home state and reachability
\end{proposition}

@@

\begin{itemize}
\item Let us choose $J=S= \downarrow S$ and $I=\uparrow x$ then resilience is equivalent to coverability of $x$ from $s$.
%\item $J=\{s\}$ and $I=\{x\}$ this particular resilience is reachability of $x$ from $s$.

\item $t$-liveness  can be expressed as the following formula
$ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ 
where
$U_t=\uparrow \pred(t)$
is the upward closure of the preconditions to use transition $t$.  
The problem reduces itself to the resilience problem
where $J=S$ and $I = \uparrow \pred(t)$.

@@

%	\mathieu{Ne marche pas (un des deux sens est invalide imho)}
%  \textcolor{red}{  Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$. Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$. Let us apply this property to LCS with $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$. This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
% Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.}
%	\mathieu{Ne marche pas (un des deux sens est invalide imho)}
\end{proof}

@@

\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
{\sc bounded resilience} is decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
\end{theorem}

We may immediately generalyse this last result by strenghthenin to \emph{unbounded} resilience. The proof is essentially the same than the previous one.

\begin{corollary}\label{postcomputable}
{\sc resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
\end{corollary}

\begin{proof}
From Fact~\ref{stop condition}, there exists $k \in \N$ such that
$\post^*(I) = I^{k_m}$. We can compute this $k_m$ by iteratively computing
$I^{k+1}$ from $I^k$, checking $I^{k+1}=I^k$, 
returning $k_m$ if that is the case.
Then, because $k_m$-resilience is decidable, 
checking $\uparrow post^*(s) \cap J \subseteq I^k = \pred^*(I)$ is,
and resilience is decidable.
% \textcolor{red}{à faire Mathieu}
\end{proof}

%\"Ozkan
\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]
For finite-branching WSTS with strong upward compatibility, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e. given a state $s$ of a WSTS
% in the regarded class 
with strong upward compatibility 
and a decidable downward-closed set $J$, it can be decided whether $\exists s' \in J ~ s \to^* s'$. 
\end{proposition}

The characterization is used to show that Petri nets are $\post^*$-effective. It is well-known that Petri nets constitute WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}). Hence Resilience is decidable for Petri nets and VASS.
However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.

@@

\begin{proof}
Let us show that Reset VASS don't enjoy the property that $\uparrow \post^*$ is computable.
Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
%in counter $i$ 
by examining if there is %some vector $v$ in the basis such that $v(i)=0$
$0$ in the basis%
. But reachability of $0$  %in a counter $i$ 
is undecidable for Reset VASS. 
% sont WSTS with strong compatibility. et l'accessibilité est undecidable.
%	$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
\end{proof}

@@

remarque si on enlève la strong monotony mais en gardant with effective basis of $\uparrow post^*$:

@@

\begin{proof}
leur algorithme termine mais pas correct car l'intersection de post et J car alors le lemmme xxxx est faux (contre-exemple)
 \textcolor{red}{à faire Mathieu}

\mathieu{En train de chercher un contre exemple pour le corollaire du Lemme~\ref{Lemma intersection}.}

\mathieu{

@@

\begin{proposition}(à prouver)
{\sc bounded resilience} is undecidable for WSTS with upward compatibility withd an effective basis of $post^*$.
\end{proposition}

@@

\begin{theorem}
{\sc Resilience} is decidable for (not necessarily strong) WSTS with effective $\uparrow$ $\post^*$ basis.
\end{theorem}

\begin{proof}
Let $B_p$ be a basis of $\uparrow \post^*(s)$, $B_I$ a basis of $I$
and $J$ decidable downward-closed.
By applying Lemma~\ref{Lemma intersection} twice, we obtain

$$ post^*(s) \cap J \subseteq \pred^*(I) \equiv B_p \cap J \subseteq \pred^*(I)$$

Since $B_p$ is finite and $J$ is decidable, we can directly compute $ B_p \cap J$.
% For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
We can compute a basis of $\pred^*(I)$ from $B_I$, and hence check that $B_p \cap J \setminus \pred^*(I) = \emptyset$. 
\end{proof}

>>>>

\problemx{$k$-reachable-resilience problem for WSTS (kRRP)}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
{ $\Bad \cap \post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
%

Analogously, we formulate bounded and general variations of the problem.

\problemx{bounded-reachable-resilience problem for WSTS (BRRP)}
{A state $s$ of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-reachable-resilient. ?\newline}

\problemx{Rachable-resilience problem for WSTS (RRP)}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
{$ ~ \forall s' \in \Bad ~ (s \rightarrow^* s') \implies \exists s'' \in \Safe ~ s' \rightarrow^{*} s''$ ?\newline}

@@

\iffalse

@@

\fi

\begin{proposition}\label{reductions}
Coverability and $t$-liveness are reduceable to the reachable resilience problem.
%	home state and reachability
\end{proposition}

@@

\begin{itemize}
\item Let us choose $\Bad=S= \downarrow S$ and $\Safe=\uparrow x$ then reachable resilience is equivalent to coverability of $x$ from $s$.

\item $t$-liveness  can be expressed as the following formula
$ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ 
where
$U_t=\uparrow \pred(t)$
is the upward closure of the preconditions to use transition $t$.  
The problem reduces itself to the reachable resilience problem
where $\Bad=S$ and $\Safe = \uparrow \pred(t)$.

@@

@@

\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
{\sc bounded-reachable-resilience} and {\sc $k$-reachable-resilience} are decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow post^*$ is computable.
\end{theorem}

We may immediately generalyse this last result by strengthening to \emph{unbounded} resilience. The proof is essentially the same than the previous one.

\begin{corollary}\label{postcomputable}
{\sc reachable-resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow post^*$ is computable.
\end{corollary}

\begin{proof}
From Fact~\ref{stop condition}, there exists $k \in \N$ such that
$\pred^*(\Safe) =  \Safe^{k_m}$. We can compute this $k_m$ by iteratively computing
$\Safe^{k+1}$ from $\Safe^k$, checking $\Safe^{k+1}=\Safe^k$, 
returning $k_m$ if that is the case.
Then, because $k_m$-reachable-resilience is decidable, 
checking $\uparrow post^*(s) \cap \Bad \subseteq \Safe^k = \pred^*(\Safe)$ is,
and reachable-resilience is decidable.
\end{proof}

%\"Ozkan
\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]
For finite-branching WSTS with strong upward compatibility, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e. given a state $s$ of a WSTS
% in the regarded class 
with strong upward compatibility 
and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. 
\end{proposition}

The characterization is used to show that Petri nets are $\post^*$-effective. It is well-known that Petri nets constitute WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}). Hence reachable-resilience is decidable for Petri nets and VASS.
However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.

@@

\begin{proof}
Let us show that Reset VASS don't enjoy the property that $\uparrow \post^*$ is computable.
Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
%in counter $i$ 
by examining if there is %some vector $v$ in the basis such that $v(i)=0$
$0$ in the basis%
. But reachability of $0$  %in a counter $i$ 
is undecidable for Reset VASS. 
% sont WSTS with strong compatibility. et l'accessibilité est undecidable.
%	$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
\end{proof}

@@

\mathieu{remarque si on enlève la strong monotony mais en gardant with effective basis of $\uparrow post^*$:}

@@

\begin{proof}
leur algorithme termine mais pas correct car l'intersection de post et J car alors le lemmme xxxx est faux (contre-exemple)
\mathieu{

@@

\begin{lemma}\label{Lemma intersection}
Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
Then $A \cap D \subseteq U$  iff $ (\uparrow  A) \cap D \subseteq U$.
\end{lemma}

\begin{proof}
Let us suppose that $A \cap D \subseteq U$. Then ${\uparrow (A \cap D)} \subseteq {\uparrow U} = U$.
Let us show that $({\uparrow A}) \cap D \subseteq {\uparrow (A \cap D)}$.
Let $x \in ({\uparrow A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
Hence $a \in A \cap D$ and then $x \in { \uparrow (A \cap D)}$.
In the other direction,
since $A \subseteq {\uparrow A}$, the inclusion
$({\uparrow  A}) \cap D \subseteq U$ implies
$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$.
\end{proof}

\begin{theorem}
{\sc Reachable-resilience} is decidable for (not necessarily strong) WSTS with effective 
$\uparrow$ $\post^*$ basis.
\end{theorem}

\begin{proof}
Let $B_p$ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
and $\Bad$ decidable downward-closed.
By applying Lemma~\ref{Lemma intersection} twice, we obtain

$$ post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \equiv B_p \cap \Bad \subseteq \pred^*(\Safe)$$

Since $B_p$ is finite and $\Bad$ is decidable, we can directly compute $ B_p \cap \Bad$.
% For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
We can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, and hence check that $B_p \cap \Bad \setminus \pred^*(\Safe) = \emptyset$. 
\end{proof}

\begin{proposition}
\textcolor{red}{
{\sc bounded resilience} is undecidable for WSTS with upward compatibility with an effective basis of $post^*$.
CONJECTURE
}
\end{proposition}

@@
